\bf Looking at Non-Linear Dimension Reductions as Models in the Data Space

### 2024-06-23 paper.qmd
re-organised the four parameter diagnostic plots

keywords: [high-dimensional data, dimension reduction, hexagonal binning, low-dimensional manifold, tour, data vizualization, model in the data space]
keywords-formatted: [high-dimensional data, dimension reduction, hexagonal binning, low-dimensional manifold, tour, data vizualization, model in the data space]

@@

$C_k^{(p)} \equiv (C_{kx_1}, ..., C_{kx_p})$ $p$-D mappings of 2D hexagonal bin centroids of the $k^{th}$ hexagon

@@

With our goal being to make a representation of this \gD{}    layout that can be lifted into high-dimensional space, the layout needs to be augmented to include neighbour information. A simple approach would be to triangulate the points and add edges. A more stable approach is to first hexagonally bin the data, reducing it from $n$ to $m\leq n$ observations, and connect the bin centroids. This process serves to reduce some noisiness in the resulting surface shown in \pD{}. The steps in this process are shown in @fig-NLDR-scurve, and documented below.

@@

<!--Full hexagonal grid with UMAP data-->

@@

## Hexagonal binning to have regular hexagons
hb_obj_scurve <- hex_binning(
  data = umap_scurve_scaled, 
  bin1 = num_bins_x_scurve, 
  r2 = r2)

@@

<!-- To define a hexagonal grid across the \kD{} space, with hexagons of fixed height ($a_1$), width ($a_2$),  it is necessary to determine how many hexagons should be represented along each axis in the \kD{} space.

@@

<!--
Define a hexagonal grid across the \kD{} space, using hexagons with fixed height ($a_2$), width ($a_1$). Each of the \kD{} points will belong to a hexagonal bin. That is, for each $y \in \mathbfit{Y}$, we can (uniquely) identify the hexagon that the point belongs to. This identification is done finding the nearest bin centroid for the \kD{} points by considering the \kD{} Euclidean distance.

@@

<!--We deliberately separate out the creation of the hexagonal grid from the mapping of points on the grid.-->

@@

The number of hexagonal bins in the hexagonal grid has a considerable impact on the construction of the \gD{} model, serving as the initial step. The chosen total number of bins must effectively capture the structure of the NLDR data. If the number of bins is too low, the model may not be able to capture the structure of the NLDR data effectively (see @fig-bins-scurve (a)), while if there are too many bins, it may result in over-fitting the individual points of the NLDR data (see @fig-bins-scurve (c)). Therefore, it is important to determine an effective number of bins to construct a reasonable model (see @fig-bins-scurve (b)).

@@

```{r}
## Hexagonal binning to have regular hexagons
hb_obj_scurve1 <- hex_binning(
  data = umap_scurve_scaled, 
  bin1 = 7, 
  r2 = r2,
  q = 0.05)

@@

```{r}
## Hexagonal binning to have regular hexagons
hb_obj_scurve2 <- hex_binning(
  data = umap_scurve_scaled, 
  bin1 = 10, 
  r2 = r2,
  q = 0.05)

@@

```{r}
## Hexagonal binning to have regular hexagons
hb_obj_scurve3 <- hex_binning(
  data = umap_scurve_scaled, 
  bin1 = 14, 
  r2 = r2,
  q = 0.05)

@@

mse_scurve_b <- ggplot(error_scurve, 
                     aes(x = b, 
                         y = log(MSE))) + 
  geom_point(size = 1) +
  geom_line() + 
  geom_vline(xintercept = 190, linetype="solid", 
             color = "black", linewidth=0.8, alpha = 0.5) +
  ylab("log(MSE)") +
  xlab("b") +
  theme(axis.text.x = element_text(size = 10),
        axis.text.y = element_text(size = 10),
        axis.title.x = element_text(size = 10),
        axis.title.y = element_text(size = 10, angle = 90))
```

```{r}
#| fig-cap: "MSE from UMAP applied to S-curve dataset with diffferent $b$ choices. What is the effective $b$ to create the model? The residual plot have a steep slope at the beginning, indicating that a smaller $b$ causes a larger amount of MSE. Then, the slope gradually declines or level off, indicating that a higher $b$ generates a smaller MSE. Using the elbow method, it was observed that when the $b$ is set to $190$, the lowest MSE occurred."
#| label: fig-mse-scurve-b
#| fig-pos: H
#| out-height: 50%

@@

Once setting up the hexagonal grid with an appropriate number of bins, some hexagonal bins may have few or no data points within them (see @fig-bins-scurve (b)). To ensure comprehensive coverage of the NLDR data, it is necessary to select hexagonal bins with a considerable number of data points. This involves calculating the number of points within each hexagon. Then, the standard count is computed by dividing the number of points within each hexagon by the maximum number of points in the grid. Next, bins with a standard count less than a benchmark value are removed (see @fig-lwd-scurve (a)). There is no specific rule for selecting a benchmark value. However, the following steps can help determine a suitable value for removing low-density hexagons:

@@

```{r}
cell_count_scurve <- ggplot(df_bin_centroids_scurve, 
                            aes(x = reorder(as.factor(hexID), 
                                            -std_counts), 
                                y = std_counts)) + 
  geom_quasirandom() + 
  xlab("hexagonal id") + 
  ylab("standardized bin count") +
  geom_hline(yintercept = benchmark1, color = "black", linewidth=0.8, alpha = 0.5) +
  theme(axis.title.x = element_text(size = 10),
        axis.title.y = element_text(size = 10, angle = 90),
        axis.text.x = element_text(size = 7, angle = 90),
        axis.text.y = element_text(size = 10))
```

```{r}
#| echo: false
#| fig-cap: "Distribution of standardize counts by hexagons."
#| label: fig-stdcts-scurve
#| out-height: 50%
#| fig-pos: H

@@

mse_scurve_lwd <- ggplot(error_rm_scurve, 
                     aes(x = benchmark_rm_lwd, 
                         y = log(MSE))) + 
  geom_point(size = 1) +
  geom_line() + 
  geom_vline(xintercept = benchmark1, linetype="solid", 
             color = "black", linewidth=0.8, alpha = 0.5) +
  geom_text(data = benchmark_label_df, 
            aes(x=x, 
                label= label, 
                y=y), 
            colour="black", 
            angle=90) +
  ylab("log(MSE)") +
  xlab("benchmark value") +
  theme(axis.text.x = element_text(size = 10),
        axis.text.y = element_text(size = 10),
        axis.title.x = element_text(size = 10),
        axis.title.y = element_text(size = 10, angle = 90))
```

```{r}
#| fig-cap: "MSE from UMAP applied to S-curve dataset with diffferent bechmark choices. What is the effective bechmark value to remove low density hexagons? The residual plot have a steep slope at the end, indicating that a smaller bechmark value causes a small amount of MSE. Then, the slope gradually increases or level up, indicating that a higher bechmark value generates a higher MSE. Using the elbow method, it was observed that when the bechmark value is set to $0.242$, the lowest MSE occurred."
#| label: fig-mse-scurve-lwd
#| fig-pos: H
#| out-height: 50%

@@

## To draw the distance distribution
distance_scurve$group <- "1"
distance_scurve_plot <- ggplot(distance_scurve, 
                        aes(x = group, 
                            y = distance)) + 
  geom_quasirandom() + 
  ylim(0, max(unlist(distance_scurve$distance))+ 0.5) + 
  coord_flip() + 
  geom_hline(yintercept = benchmark_scurve, linetype="solid", 
             color = "black", linewidth=0.8, alpha = 0.5) +
  geom_text(data = benchmark_label_df, 
            aes(x=x, 
                label= label, 
                y=y), 
            colour="black", 
            angle=90) +
  ylab(expression(d^{(2)})) +
  theme(axis.text.x = element_text(size = 10),
        axis.title.x = element_text(size = 10))
```

```{r}
#| echo: false
#| fig-cap: "Distribution of $2\\text{-}D$ Euclidean distances between bin centroids of the triangular mesh generated with S-curve UMAP data."
#| label: fig-dist-scurve
#| out-height: 50%
#| fig-pos: H

distance_scurve_plot
```

>>>>

keywords: [high-dimensional data, dimension reduction, hexagon binning, low-dimensional manifold, tour, data vizualization, model in the data space]
keywords-formatted: [high-dimensional data, dimension reduction, hexagon binning, low-dimensional manifold, tour, data vizualization, model in the data space]

@@

$C_k^{(p)} \equiv (C_{kx_1}, ..., C_{kx_p})$ $p$-D mappings of 2D hexagon bin centroids of the $k^{th}$ hexagon

@@

With our goal being to make a representation of this \gD{}    layout that can be lifted into high-dimensional space, the layout needs to be augmented to include neighbour information. A simple approach would be to triangulate the points and add edges. A more stable approach is to first hexagonly bin the data, reducing it from $n$ to $m\leq n$ observations, and connect the bin centroids. This process serves to reduce some noisiness in the resulting surface shown in \pD{}. The steps in this process are shown in @fig-NLDR-scurve, and documented below.

@@

<!--Full hexagon grid with UMAP data-->

@@

## hexagon binning to have regular hexagons
hb_obj_scurve <- hex_binning(
  data = umap_scurve_scaled, 
  bin1 = num_bins_x_scurve, 
  r2 = r2)

@@

<!-- To define a hexagon grid across the \kD{} space, with hexagons of fixed height ($a_1$), width ($a_2$),  it is necessary to determine how many hexagons should be represented along each axis in the \kD{} space.

@@

<!--
Define a hexagon grid across the \kD{} space, using hexagons with fixed height ($a_2$), width ($a_1$). Each of the \kD{} points will belong to a hexagon bin. That is, for each $y \in \mathbfit{Y}$, we can (uniquely) identify the hexagon that the point belongs to. This identification is done finding the nearest bin centroid for the \kD{} points by considering the \kD{} Euclidean distance.

@@

<!--We deliberately separate out the creation of the hexagon grid from the mapping of points on the grid.-->

@@

The number of hexagon bins in the hexagon grid has a considerable impact on the construction of the \gD{} model, serving as the initial step. The chosen total number of bins must effectively capture the structure of the NLDR data. If the number of bins is too low, the model may not be able to capture the structure of the NLDR data effectively (see @fig-bins-scurve (a)), while if there are too many bins, it may result in over-fitting the individual points of the NLDR data (see @fig-bins-scurve (c)). Therefore, it is important to determine an effective number of bins to construct a reasonable model (see @fig-bins-scurve (b)).

@@

```{r}
## hexagon binning to have regular hexagons
hb_obj_scurve1 <- hex_binning(
  data = umap_scurve_scaled, 
  bin1 = 7, 
  r2 = r2,
  q = 0.05)

@@

```{r}
## hexagon binning to have regular hexagons
hb_obj_scurve2 <- hex_binning(
  data = umap_scurve_scaled, 
  bin1 = 10, 
  r2 = r2,
  q = 0.05)

@@

```{r}
## hexagon binning to have regular hexagons
hb_obj_scurve3 <- hex_binning(
  data = umap_scurve_scaled, 
  bin1 = 14, 
  r2 = r2,
  q = 0.05)

@@

mse_scurve_b <- ggplot(error_scurve, 
                     aes(x = b, 
                         y = log(MSE))) + 
  geom_point(size = 1) +
  geom_line() + 
  geom_vline(xintercept = 190, linetype="solid", 
             color = "black", linewidth=0.8, alpha = 0.5) +
  ylab("log(MSE)") +
  xlab("b") +
  ggtitle("(a)") +
  theme(aspect.ratio = 0.75,
        axis.text.x = element_text(size = 10),
        axis.text.y = element_text(size = 10),
        axis.title.x = element_text(size = 10),
        axis.title.y = element_text(size = 10, 
            angle=90),
        title = element_text(size = 36))
```

```{r}
#| fig-cap: "MSE from UMAP applied to S-curve dataset with diffferent $b$ choices. What is the effective $b$ to create the model? The residual plot have a steep slope at the beginning, indicating that a smaller $b$ causes a larger amount of MSE. Then, the slope gradually declines or level off, indicating that a higher $b$ generates a smaller MSE. Using the elbow method, it was observed that when the $b$ is set to $190$, the lowest MSE occurred."
#| label: fig-mse-scurve-b
#| fig-pos: H
#| out-height: 50%
#| eval: false

@@

Once setting up the hexagon grid with an appropriate number of bins, some hexagon bins may have few or no data points within them (see @fig-bins-scurve (b)). To ensure comprehensive coverage of the NLDR data, it is necessary to select hexagon bins with a considerable number of data points. This involves calculating the number of points within each hexagon. Then, the standard count is computed by dividing the number of points within each hexagon by the maximum number of points in the grid. Next, bins with a standard count less than a benchmark value are removed (see @fig-lwd-scurve (a)). There is no specific rule for selecting a benchmark value. However, the following steps can help determine a suitable value for removing low-density hexagons:

@@

```{r}
cell_count_scurve <- ggplot(df_bin_centroids_scurve, 
                            aes(x = reorder(as.factor(hexID), 
                                            -std_counts), 
                                y = std_counts)) + 
  geom_quasirandom() + 
  geom_hline(yintercept = benchmark1, color = "black", linewidth=0.8, alpha = 0.5) +
  xlab("hexagon id (re-ordered)") + 
  ylab("standardized bin count") +
  ggtitle("(b)") +
  theme(aspect.ratio = 0.75,
        axis.title.x = element_text(size = 10),
        axis.title.y = element_text(size = 10, angle = 90),
        #axis.text.x = element_text(size = 7, angle = 90),
        axis.text.x = element_blank(),
        axis.text.y = element_text(size = 10),
        title = element_text(size = 36))
```

```{r}
#| echo: false
#| fig-cap: "Distribution of standardize counts by hexagons."
#| label: fig-stdcts-scurve
#| out-height: 50%
#| fig-pos: H
#| eval: false

@@

mse_scurve_lwd <- ggplot(error_rm_scurve, 
                     aes(x = benchmark_rm_lwd, 
                         y = log(MSE))) + 
  geom_point(size = 1) +
  geom_line() + 
  geom_vline(xintercept = benchmark1, linetype="solid", 
             color = "black", linewidth=0.8, alpha = 0.5) +
  geom_text(data = benchmark_label_df, 
            aes(x=x, 
                label= label, 
                y=y), 
            colour="black", 
            angle=90) +
  scale_x_continuous("standardized bin count", 
         transform = "reverse") + 
  ylab("log(MSE)") +
  ggtitle("(c)") +
  theme(aspect.ratio = 0.75,
        axis.text.x = element_text(size = 10),
        axis.text.y = element_text(size = 10),
        axis.title.x = element_text(size = 10),
        axis.title.y = element_text(size = 10, 
            angle=90),
        title = element_text(size = 36))
```

```{r}
#| fig-cap: "MSE from UMAP applied to S-curve dataset with diffferent bechmark choices. What is the effective bechmark value to remove low density hexagons? The residual plot have a steep slope at the end, indicating that a smaller bechmark value causes a small amount of MSE. Then, the slope gradually increases or level up, indicating that a higher bechmark value generates a higher MSE. Using the elbow method, it was observed that when the bechmark value is set to $0.242$, the lowest MSE occurred."
#| label: fig-mse-scurve-lwd
#| fig-pos: H
#| out-height: 50%
#| eval: false

@@

## To draw the distance distribution
distance_scurve$group <- "1"
distance_scurve_plot <- ggplot(distance_scurve, 
                        aes(x = group, 
                            y = distance)) + 
  geom_quasirandom() + 
  ylim(0, max(unlist(distance_scurve$distance))+ 0.5) + 
  coord_flip() + 
  geom_hline(yintercept = benchmark_scurve,
     linetype="solid", 
     color = "black", linewidth=0.8, alpha = 0.5) +
  geom_text(data = benchmark_label_df, 
            aes(x=x, 
                label= label, 
                y=y), 
            colour="black", 
            angle=90) +
  ylab(expression(d^{(2)})) +
  ggtitle("(d)") +
  theme(aspect.ratio = 0.75,
        axis.text.x = element_text(size = 10),
        axis.title.x = element_text(size = 10),
        title = element_text(size = 36))
```

```{r}
#| echo: false
#| fig-cap: "Various plots to help assess best number of bins, low density bin and large edge removal. Both (a) and (c) show MSE, against number of bins and standardised count. A good benchmark value for these parameters is when the MSE drops and then flattens out."
# Distribution of $2\\text{-}D$ Euclidean distances between bin centroids of the triangular mesh generated with S-curve UMAP data.
#| label: fig-dist-scurve
#| out-width: 100%
#| fig-width: 8
#| fig-height: 6
#| fig-pos: H

mse_scurve_b + cell_count_scurve + 
mse_scurve_lwd + 
distance_scurve_plot + plot_layout(ncol=2)
```

### 2024-06-23 paper.tex
re-organised the four parameter diagnostic plots

\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same} % disable monospaced font for URLs
\hypersetup{
  pdftitle={Looking at Non-Linear Dimension Reductions as Models in the Data Space},
  pdfauthor={Jayani P.G. Lakshika; Dianne Cook; Paul Harrison; Michael Lydeamore; Thiyanga S. Talagala},
  pdfkeywords={high-dimensional data, dimension reduction, hexagonal
binning, low-dimensional manifold, tour, data vizualization, model in
the data space},
  colorlinks=true,
  linkcolor={blue},
  filecolor={Maroon},
  citecolor={Blue},
  urlcolor={Blue},
  pdfcreator={LaTeX via pandoc}}

@@

\noindent%
{\it Keywords:} high-dimensional data, dimension reduction, hexagonal
binning, low-dimensional manifold, tour, data vizualization, model in
the data space
\vfill

@@

With our goal being to make a representation of this \gD{} layout that
can be lifted into high-dimensional space, the layout needs to be
augmented to include neighbour information. A simple approach would be
to triangulate the points and add edges. A more stable approach is to
first hexagonally bin the data, reducing it from \(n\) to \(m\leq n\)
observations, and connect the bin centroids. This process serves to
reduce some noisiness in the resulting surface shown in \pD{}. The steps
in this process are shown in Figure~\ref{fig-NLDR-scurve}, and
documented below.

@@

The number of hexagonal bins in the hexagonal grid has a considerable
impact on the construction of the \gD{} model, serving as the initial
step. The chosen total number of bins must effectively capture the
structure of the NLDR data. If the number of bins is too low, the model
may not be able to capture the structure of the NLDR data effectively
(see Figure~\ref{fig-bins-scurve} (a)), while if there are too many
bins, it may result in over-fitting the individual points of the NLDR
data (see Figure~\ref{fig-bins-scurve} (c)). Therefore, it is important
to determine an effective number of bins to construct a reasonable model
(see Figure~\ref{fig-bins-scurve} (b)).

@@

To determine the effective \(b\), candidate values are selected based on
the range between the minimum and approximate maximum \(b_1\), because
\(b_2\) is computed from \(b_1\). The minimum \(b_1\) is set to \(2\),
while the maximum number is estimated by taking the square root of
\(\frac{n}{2}\). The analysis evaluates the MSE across varying \(b\)
within this range, covering the minimum to maximum values along both
axes (see Figure~\ref{fig-mse-scurve-b}).

\includegraphics[width=1\textwidth,height=0.5\textheight]{paper_files/figure-pdf/fig-mse-scurve-b-1.pdf}

\caption{\label{fig-mse-scurve-b}MSE from UMAP applied to S-curve
dataset with diffferent \(b\) choices. What is the effective \(b\) to
create the model? The residual plot have a steep slope at the beginning,
indicating that a smaller \(b\) causes a larger amount of MSE. Then, the
slope gradually declines or level off, indicating that a higher \(b\)
generates a smaller MSE. Using the elbow method, it was observed that
when the \(b\) is set to \(190\), the lowest MSE occurred.}

@@

Once setting up the hexagonal grid with an appropriate number of bins,
some hexagonal bins may have few or no data points within them (see
Figure~\ref{fig-bins-scurve} (b)). To ensure comprehensive coverage of
the NLDR data, it is necessary to select hexagonal bins with a
considerable number of data points. This involves calculating the number
of points within each hexagon. Then, the standard count is computed by
dividing the number of points within each hexagon by the maximum number
of points in the grid. Next, bins with a standard count less than a
benchmark value are removed (see Figure~\ref{fig-lwd-scurve} (a)). There
is no specific rule for selecting a benchmark value. However, the
following steps can help determine a suitable value for removing
low-density hexagons:

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item
  Plot the distribution of the standardized counts (see
  Figure~\ref{fig-stdcts-scurve}).
\item
  Examine the distribution of counts.
\item
  Select the first quantile value if the distribution is skewed.
\end{enumerate}

\includegraphics[width=1\textwidth,height=0.5\textheight]{paper_files/figure-pdf/fig-stdcts-scurve-1.pdf}

\caption{\label{fig-stdcts-scurve}Distribution of standardize counts by
hexagons.}

The benchmark value for removing low-density hexagons ranges between
\(0\) and \(1\). When analyzing how these benchmark values influence
model performance, it's essential to observe the change in MSE as the
benchmark value increases (see Figure~\ref{fig-mse-scurve-lwd}). The MSE
shows a gradual increase as the benchmark value progresses from \(0\) to
\(1\). Evaluating this rate of increase is important. If the increment
is not considerable, the decision might lean towards retaining
low-density hexagons.

\includegraphics[width=1\textwidth,height=0.5\textheight]{paper_files/figure-pdf/fig-mse-scurve-lwd-1.pdf}

\caption{\label{fig-mse-scurve-lwd}MSE from UMAP applied to S-curve
dataset with diffferent bechmark choices. What is the effective bechmark
value to remove low density hexagons? The residual plot have a steep
slope at the end, indicating that a smaller bechmark value causes a
small amount of MSE. Then, the slope gradually increases or level up,
indicating that a higher bechmark value generates a higher MSE. Using
the elbow method, it was observed that when the bechmark value is set to
\(0.242\), the lowest MSE occurred.}

@@

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item
  Plot the distribution of the 2D Euclidean distances (see
  Figure~\ref{fig-dist-scurve}).
\item
  Identify the first largest difference between consecutive distance
  values.
\item
  Take the distance value corresponding to this difference as the
  benchmark value.
\end{enumerate}

@@

\includegraphics[width=1\textwidth,height=0.5\textheight]{paper_files/figure-pdf/fig-dist-scurve-1.pdf}

@@

\caption{\label{fig-dist-scurve}Distribution of \(2\text{-}D\) Euclidean
distances between bin centroids of the triangular mesh generated with
S-curve UMAP data.}

>>>>

\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same} % disable monospaced font for URLs
\hypersetup{
  pdftitle={Looking at Non-Linear Dimension Reductions as Models in the Data Space},
  pdfauthor={Jayani P.G. Lakshika; Dianne Cook; Paul Harrison; Michael Lydeamore; Thiyanga S. Talagala},
  pdfkeywords={high-dimensional data, dimension reduction, hexagon
binning, low-dimensional manifold, tour, data vizualization, model in
the data space},
  colorlinks=true,
  linkcolor={blue},
  filecolor={Maroon},
  citecolor={Blue},
  urlcolor={Blue},
  pdfcreator={LaTeX via pandoc}}

@@

\noindent%
{\it Keywords:} high-dimensional data, dimension reduction, hexagon
binning, low-dimensional manifold, tour, data vizualization, model in
the data space
\vfill

@@

With our goal being to make a representation of this \gD{} layout that
can be lifted into high-dimensional space, the layout needs to be
augmented to include neighbour information. A simple approach would be
to triangulate the points and add edges. A more stable approach is to
first hexagonly bin the data, reducing it from \(n\) to \(m\leq n\)
observations, and connect the bin centroids. This process serves to
reduce some noisiness in the resulting surface shown in \pD{}. The steps
in this process are shown in Figure~\ref{fig-NLDR-scurve}, and
documented below.

@@

The number of hexagon bins in the hexagon grid has a considerable impact
on the construction of the \gD{} model, serving as the initial step. The
chosen total number of bins must effectively capture the structure of
the NLDR data. If the number of bins is too low, the model may not be
able to capture the structure of the NLDR data effectively (see
Figure~\ref{fig-bins-scurve} (a)), while if there are too many bins, it
may result in over-fitting the individual points of the NLDR data (see
Figure~\ref{fig-bins-scurve} (c)). Therefore, it is important to
determine an effective number of bins to construct a reasonable model
(see Figure~\ref{fig-bins-scurve} (b)).

@@

To determine the effective \(b\), candidate values are selected based on
the range between the minimum and approximate maximum \(b_1\), because
\(b_2\) is computed from \(b_1\). The minimum \(b_1\) is set to \(2\),
while the maximum number is estimated by taking the square root of
\(\frac{n}{2}\). The analysis evaluates the MSE across varying \(b\)
within this range, covering the minimum to maximum values along both
axes (see \textbf{?@fig-mse-scurve-b}).

@@

Once setting up the hexagon grid with an appropriate number of bins,
some hexagon bins may have few or no data points within them (see
Figure~\ref{fig-bins-scurve} (b)). To ensure comprehensive coverage of
the NLDR data, it is necessary to select hexagon bins with a
considerable number of data points. This involves calculating the number
of points within each hexagon. Then, the standard count is computed by
dividing the number of points within each hexagon by the maximum number
of points in the grid. Next, bins with a standard count less than a
benchmark value are removed (see Figure~\ref{fig-lwd-scurve} (a)). There
is no specific rule for selecting a benchmark value. However, the
following steps can help determine a suitable value for removing
low-density hexagons:

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item
  Plot the distribution of the standardized counts (see
  \textbf{?@fig-stdcts-scurve}).
\item
  Examine the distribution of counts.
\item
  Select the first quantile value if the distribution is skewed.
\end{enumerate}

The benchmark value for removing low-density hexagons ranges between
\(0\) and \(1\). When analyzing how these benchmark values influence
model performance, it's essential to observe the change in MSE as the
benchmark value increases (see \textbf{?@fig-mse-scurve-lwd}). The MSE
shows a gradual increase as the benchmark value progresses from \(0\) to
\(1\). Evaluating this rate of increase is important. If the increment
is not considerable, the decision might lean towards retaining
low-density hexagons.

@@

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item
  Plot the distribution of the 2D Euclidean distances (see
  \textbf{?@fig-dist-scurve}).
\item
  Identify the first largest difference between consecutive distance
  values.
\item
  Take the distance value corresponding to this difference as the
  benchmark value.
\end{enumerate}

@@

{\centering \includegraphics[width=1\textwidth,height=\textheight]{paper_files/figure-pdf/unnamed-chunk-31-1.pdf}

@@

\caption{Various plots to help assess best number of bins, low density
bin and large edge removal. Both (a) and (c) show MSE, against number of
bins and standardised count. A good benchmark value for these parameters
is when the MSE drops and then flattens out.}
